\section{Discussion}\label{s:discussion}

%%A “limitations” section, as the name implies, describes scenarios where the
%%proposed solution may not work well. Although a “discussion” section could also
%highlight limitations of the proposed work, it focuses on analyzing the
%implications of the proposed work for current and future research.

\subsection{Design Limitations}
The implemented detection mitigations are system wide; this provides the best level of transparency, even in the case of an attempt 
to perform these detection checks on a different process. The downside of this approach is that it will also interfere with normal, 
genuine processes that expect certain data or values to be present, like the VM support programs many commercial hypervisor vendors use, 
which might store crucial data in locations blocked by these transparency features. As well as Windows processes that expect no kernel modifications. 
As an addition to the transparency mode in HyperDbg, a two stage transparency, with a semi-transparent mode added, would improve the usability and reliability of the hypervisor. 
In this system, a select number of processes are allowed to bypass all transparent mitigations as well as not intercepting the MSR accesses that could cause instability. 
As a trade-off, this semi-transparent mode would allow malware to more easily run successful hypervisor detections and avoid full analysis, by emulating the appearance of one 
of these genuine, whitelisted processes.

The system call interception of the mitigation design is based on filtering the system calls based on their system call numbers~\ref{tab:syscalls} and thus the mitigations 
only work if deployed on a guest system that runs the correct Windows version, as Microsoft is known to change the system call numbers between versions 
frequently~\cite{j00ruSyscalls}. In an implementation of the system call interception, an ability to obtain these system call numbers is needed for correct operation.
Within HyperDbg, obtaining the current Windows version of the guest is possible, and a feature to get the system calls of the guest OS version was also implemented, but not as a part of this thesis.

\subsection{Not Full Transparency}
This thesis only covered 3 general hypervisor detection vectors, there are numerous other ways any user-space process could detect the virtualized environment, 
and a fully transparent and reliable hypervisor is nearly impossible to create. Deploying HyperDbg with the transparency features from this thesis does not guarantee any true 
transparency, and it can remain stealthy if the malware or any other binary analyzed only uses the detection methods discussed in this thesis.

The implemented transparency features also leave detectable traces of their own, and a process developed with HyperDbg in mind could find these 
traces and still detect the virtualized environment of HyperDbg's hypervisor. The remaining artifacts include inconsistent return values, memory buffers 
that are not the correct size, or inconsistencies between the same data obtained from different sources, like a system call which does not get intercepted by HyperDbg.

While guaranteeing 100\% transparency is not feasible, the design created during this thesis significantly raises the bar and makes it more difficult for malware to defeat these anti-anti-hypervisor techniques.



%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../thesis"
%%% End:
